---
title: "Visualizing logistic regression results for non-technical audiences"
author: "Abby Kaplan and Keiko Cawley"
date: "9/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(ggplot2)
```



## Introduction: Just tell me "the" effect

Oftentimes stakeholders are interested in whether something has a positive or negative effect and if so, by how much. When we use a logistic regression to model a binary outcome, communicating these results is not so straightforward. The crux of the issue is due to the non-linear nature of the logistic regression curve.

To review, the formula for the logistic regression model is the following: 

$$
p =\frac{exp(\beta_0 + \beta_1X_1 + ... + \beta_nX_n)}{1 + exp(\beta_0 + \beta_1X_1 + ... + \beta_nX_n)}
$$
In the equation *p* is the probability, each of the $\beta$ coefficients represents some increase or decrease on the logistic regression curve, and each of the X~i~s are either ones or zeros which act like switches to turn the $\beta$s on or off. 

If we graph the logistic regression equation we get a sigmoid shaped curve:

```{r, echo = F, message = F}

curve <- function(x) (exp(x)/(1+(exp(x))))

set.seed(1500)
ggplot(data = data.frame(x=rnorm(10000)), mapping = aes(x = x)) +
    stat_function(fun = curve, color = "#009DD8") +
    ylim(0,1)+
    xlim(-10,10) + 
    theme(axis.ticks.x = element_blank()) +
    ylab('probability (p)') +
    xlab(label = expression(Z~'='~beta[0] + beta[1]~X[1]~'+'~'... +'~beta[n]~X[n]))

```

Two characteristics of the logistic regression's results which make the magnitude of effect difficult to convey are:

1. Where a point is on the curve above depends on what combinations of $\beta$s were "switched on" by the X~i~s. 
2. The slope of the curve is not constant: A one unit increase for different values of Z can result in vastly different probabilities *p*. Notice that the difference in probability between any two points where the slope of the curve is relatively flat, for example from x = -5 to x = -2.5, the difference is very small. On the otherhand, the difference in probability between any two points where the slope curve is steep, for example from x = 0 to x = 2.5, the difference is large. 

Due to the probability being (1) contingent on which $\beta$ parameters and (2) due to the non-constant slope means that **the magnitude of effect varies**. If the effect varies, then what is "the" effect that we report, and particularly to a stakeholder who does not know what a logistic regression is? In this talk we will explore various visualization options to present logistic regression results to non-technical audiences, and the pros and cons of these different visualizations. We will also discuss in which situations you might choose one visualization over another. 





